Models of lens candidates from Space Warps CFHTLS

### 2017-03-20 parts/intro.tex

The expected flood of new lens discoveries will need a similarly huge
modelling effort to reconstruct their mass distributions.  To prepare
for the challenge of massive-sample lens modelling,
\cite{2015MNRAS.447.2170K} developed a new modelling strategy,
implemented as the SpaghettiLens system.  The idea is to collaborate
with experienced members of the citizen-science community, who have
already participated in lens discovery through Space~Warps, as well as
several other projects involving astronomical data.  This paper
continues by applying SpaghettiLens to lens-candidates discovered
through Space~Warps.  We present results from modelling of 42 of the
59 lens candidates reported by \cite{2016MNRAS.455.1191M}.

The aim of this work is twofold.  First, one would like to identify
the most probable lens and the systems likely to be most rewarding for
future follow-up observations.  For this we introduce some
diagnostics; these

  Its purpose is twofold.  First to check if the
created models look like plausible lenses and whether the shortcomings
diagnosed in \cite{2015MNRAS.447.2170K} were fixed.  Second to
introduce diagnostic criteria that help filter out   While introducing these diagnostic criteria, we also
gain some preliminary estimations of some properties of the
lens-candidates as galaxies.

The analysis presented in this paper is somewhat limited by the fact that no source redshifts are available yet and that the redshifts of the lenses have large error bars.
Source redshifts were set to z=2 for all systems. %TODO justify z=2.

Nevertheless it is interesting to see what trends we can observe with the already available data.
One should also keep in mind that the existing pipeline allows to replace those values with more precise ones as soon as they are available and rerun this analysis without any further work involved.

% It's purpose is twofold: First, to see if
% plausible mass models can produce the overall configuration.  Second, to
% check their interpretation as lensing galaxies.

Each lens candidate is modelled in a collaborative refinement
process, where anyone interested can create a new model or modify an
existing model to try and make it better.\footnote{This is in contrast
  to the main Space~Warps project for discovering lenses, where
  volunteers in a crowd of $\gtrsim10^4$ make independent
  contributions.  Each person is presented with a random selection of
  survey-patches and invited to (in effect) vote on each.  The system
  estimates each volunteer's skill level according to test-patches
  interspersed with the real data, and weights their votes accordingly
  \citep{2016MNRAS.455.1171M}.  There is an active forum for
  volunteers, but since everyone is seeing different data samples with
  minimal overlap, the forum has little if any influence on votes.}
The result represents a consensus among contributors, as to the best
that could be achieved with the available data and software.

The modelling process assigns seven diagnostics to each lens
candidate, as follows.

@@

Section~\ref{sec:summary} summarises and tabulates the diagnostics in
Table~\ref{tab:models}.

@@

There is an appendix section
devoted to various technical issues relating to modelling.
\cite{2015MNRAS.447.2170K} tested the system on simulated lenses and
identified some areas for improvement.
In \S~\ref{subsec:sourcefit} we introduce fitting of the brightness
profiles of the source.  This feature has not yet been included in
SpaghettiLens, but has been carried out in post-processing for a few
especially interesting candidates.  In \S~\ref{subsec:hires} we show
that making mass maps fine-grained in the central region relieves a
tendency in the earlier work for mass to be too shallow. Then in
\S~\ref{subsec:parameter} we consider the possibility of fitting a
parametric lens model to the model ensemble; so far we have only been
successful at extracting an Einstein radius.

>>>>

The expected flood of new lens discoveries will need a similarly huge
modelling effort to reconstruct their mass distributions.  To prepare
for the challenge of massive-sample lens modelling,
\cite{2015MNRAS.447.2170K} developed a new modelling strategy,
implemented as the SpaghettiLens system.  The idea is to collaborate
with experienced members of the citizen-science community, who have
already participated in lens discovery through Space~Warps, as well as
several other projects involving astronomical data.  The system was
tested on a sample of simulated lenses from Space~Warps.

This paper continues by applying SpaghettiLens to lens-candidates
discovered through Space~Warps.  We present results from modelling of
42 of the 59 lens candidates reported by \cite{2016MNRAS.455.1191M}.
Each lens candidate was modelled in a collaborative refinement
process, where anyone interested can create a new model or modify an
existing model to try and make it better.\footnote{This is in contrast
  to the main Space~Warps project for discovering lenses, where
  volunteers in a crowd of $\gtrsim10^4$ make independent
  contributions.  Each person is presented with a random selection of
  survey-patches and invited to (in effect) vote on each.  The system
  estimates each volunteer's skill level according to test-patches
  interspersed with the real data, and weights their votes accordingly
  \citep{2016MNRAS.455.1171M}.  There is an active forum for
  volunteers, but since everyone is seeing different data samples with
  minimal overlap, the forum has little if any influence on votes.}
The result model represents a consensus among contributors, as to the
best that could be achieved with the available data and software.

We characterise each model with seven diagnostics, whose purpose is to
help identify which systems are most probably lenses, and which ones
are likely to be most rewarding for future follow-up observations.
The diagnostics are as follows.

@@

Section~\ref{sec:summary} summarises and tabulates the diagnostics in
Table~\ref{tab:models}.  Interpretation of the results is preliminary,
because the systems are candidates at this stage, not secure lenses.
Moreover the candidate-lens redshifts have large uncertainties, while
the candidate-source redshifts can only be guessed at present.
Nevertheless it is interesting to see what trends we can observe with
the already-available data.

There are three appendices devoted to various technical issues
relating to modelling.  \cite{2015MNRAS.447.2170K} tested the system
on simulated lenses and identified some areas for improvement.  In
\S~\ref{subsec:sourcefit} we introduce fitting of the brightness
profiles of the source.  This feature has not yet been included in
SpaghettiLens, but has been carried out in post-processing for a few
especially interesting candidates.  In \S~\ref{subsec:hires} we show
that making mass maps fine-grained in the central region relieves a
tendency in the earlier work for mass to be too shallow. Then in
\S~\ref{subsec:parameter} we consider the possibility of fitting a
parametric lens model to the model ensemble; so far we have only been
successful at extracting an Einstein radius.

@@
